Argumentation theory for mathematical argument\thanks{Support from the UK Engineering and Physical Sciences Research Council is acknowledged under grants EPSRC EP/K040251/2 (Corneli, Martin and Nesin), EP/J017728/2 (Murray-Rust) and EP/P017320/1 (Pease).}

### 2018-07-11 main.tex
grammar check

%%% Some complaint...
%%% https://tex.stackexchange.com/questions/257829/package-titlesec-error-entered-in-horizontal-mode-sectionintrodu%C3%87%C3%83o
%% \usepackage[calcwidth]{titlesec}
%% \newlength\mylen
%% \setlength\mylen{\dimexpr\oddsidemargin+1in+\hoffset\relax}

%% %\bfseries
%% \titleformat{\paragraph}[runin]
%%   {\sffamily\fontseries{sbc}\selectfont}
%%   {\protect\marginnote[\theparagraph]{\theparagraph}}{0em}{}{\space}

%% \renewcommand{\marginnotevadjust}{-1\baselineskip}

%% \titlespacing{\paragraph}{0.0em}{1ex}{0.5em}

@@

\begin{abstract}
To adequately model mathematical arguments the analyst must be able to
represent the mathematical objects under discussion and the
relationships between them, as well as inferences drawn about these
objects and relationships as the discourse unfolds.
We introduce a framework with these properties, which has been 
used to analyse mathematical dialogues and expository texts.
The framework can recover salient elements of discourse at, and within, the sentence
level, as well as the way mathematical content connects to form larger
argumentative structures.  We show how the framework might be used to
support computational reasoning, and argue that it provides a more
natural way to examine the process of proving theorems than do
Lamport's structured proofs.
%%
\keywords{inference anchoring theory \and mathematical practice \and mathematical argument \and structured proof}
\end{abstract}

@@

Indeed, IATC could be overlayed on top of mathematics represented in more formal languages, where an informal text can be put in full or partial correspondence with a formal treatment. Lamport's ``Temporal Logic of Actions+'' (TLA+) \cite{lamport1999specifying,lamport2014tla2} is one candidate formalism that could be used to model content-level expressions, by analogy with Lamport's own semi-formal ``structured proofs'' 
\cite{lamport1995write,lamport2012write21st}.

Unlike structured proof, IATC does not aim to reshape the way people do mathe- matics, but to model it more exactly. As such, it constitutes groundwork for a future generation of computer systems that can collaborate with mathematicians and students in a way these potential users already understand. \citet{epstein2015wanted} highlights the ``extent to which a person believes that her work experience or product has been facilitated or improved by the collaboration'' as a key evaluation metric for assessing collaborative intelligent computer systems. The key metric at this stage is more basic, namely, we are interested in the degree to which IATC can represent real-world examples of mathematical practice in a way that can make them accessible to computational reasoning. After introducing the modelling approach, we use several examples to show that IATC is indeed satisfactory in this regard.

%% \thesis{Argumentation theory can help us understand the way mathematical knowledge develops.}
%% In order to better understand the processes of mathematical proof and
%% communication, we turn to informal mathematical dialogues and
%% expository texts as our primary sources of data. Improved
%% understanding of mathematical practice can, in turn, inform the design
%% of support tools for mathematicians.

%% \paragraph{Computational tools for proof checking, automatic theorem proving, and computer algebra are well-established, however, they can only be used with formal, computationally explicit, content.}
%% Translating proofs into formal languages constitutes an unrealistic
%% demand for most mathematicians and students, and does little to address
%% their day-to-day communication needs \cite{thurston1994proof,melis2005proof}.
%% Furthermore, the existing mathematical literature is opaque to current computational systems,
%% which rely on hand-coded knowledge bases entered in formal or
%% constrained languages \cite{furse1990did,furse1994mathematics}.
%% %
%% Both data ingestion and end-user interaction could benefit from the use
%% of an expressive modelling language that is closely related to the way
%% mathematics is actually done.

%% \paragraph{Our approach to modelling mathematical arguments is inspired by the general-purpose argument modelling formalism \emph{Inference Anchoring Theory} (IAT), introduced by \citet{reed2011dialogues}.}
%% As its name suggests, IAT anchors logical inferences in discourse.
%% IAT has been applied to mediation \cite{janier2017towards}, debates
%% \cite{budzynska2014model}, and to paradoxes in ethotic argumentation
%% \cite{budzynska2013circularity}, along with other real-world
%% dialogues \cite{budzynska2013towards}.
%% %%
%% The \emph{Inference Anchoring Theory + Content} (IATC)
%% framework we introduce is based on IAT, but with several
%% significant modifications.  Most fundamentally,
%% IATC is designed to bring to the surface the structural features inherent in mathematical content.

%% \paragraph{Indeed, IATC could be overlayed on top of mathematics represented in more formal languages, where an informal text can be put in full or partial correspondence with a formal treatment.}
%% Lamport's ``Temporal Logic of Actions+'' ({TLA}$^{+}$)
%% \cite{lamport1999specifying,lamport2014tla2} is one candidate
%% formalism that could be used to model content-level expressions, by
%% analogy with Lamport's own semi-formal ``structured proofs''
%% \cite{lamport1995write,lamport2012write21st}.

%% \paragraph{Unlike structured proof, IATC does not aim to reshape the way people do mathematics, but to model it more exactly.}
%% As such, it constitutes groundwork for a future generation of computer systems that can collaborate with mathematicians and students in a way these potential users already understand.
%% \citet{epstein2015wanted} highlights the ``extent to which a person
%% believes that her work experience or product has been facilitated or
%% improved by the collaboration'' as a key evaluation metric for
%% assessing collaborative intelligent computer systems.
%% %%
%% The key metric at this stage is more basic, namely, we are interested
%% in the degree to which IATC can represent real-world examples of
%% mathematical practice in a way that can make them accessible to
%% computational reasoning.  After introducing the modelling approach, we
%% use several examples to show that IATC is indeed satisfactory in this
%% regard.

%% Because these examples form the key piece of empirical evidence, we
%% introduce them briefly here.  This this shows the scope and also the
%% limitations of our current approach.

\begin{itemize}
\item Our first example is a school-level challenge problem that was
  presented in a public lecture by the mathematician Timothy Gowers
  \cite{gowers-talk}.  The lecture aimed to motivate and contextualise
  a project, then beginning, to develop mathematical software that
  ``operate[s] in a way that closely mirrors the way human
  mathematicians operate'' \cite[p.~255]{ganesalingam2016fully}.  The
  reasoning needed to solve the challenge problem remains beyond the
  scope of the computational method that Ganesalingam and Gowers
  ultimately published, but it is both sufficiently simple and
  sufficiently realistic to introduce the practical aspects of
  working with IATC.
\item Our second example is a question posed on the online Q\&A forum
  MathOverflow, together with the ensuing dialogue.  MathOverflow is
  part of the Stack Exchange network of community question-and-answer
  websites, which is particularly popular with software developers.
  The MathOverflow sub-site is devoted to discussions about
  research-level questions in mathematics.  Such discussions are very
  different from the textbook-style proofs treated by
  \cite{ganesalingam2016fully}, and we discuss the considerations that
  such discussions would impose on computational modelling efforts.
\item \paragraph{MiniPolymath 1 through 4 were part of a series of
  experiments in collaborative online mathematics known as ``Polymath
  projects'' \cite{polymath-wiki}.}  While other projects in the series
  tackled novel research, the problems in the MiniPolymath subseries
  were drawn from the Mathematical Olympiad, a premier competition for
  pre-college students.
%%
Six problems are given, and the examination takes place over
two days with three problems to be solved each day.
Whereas individual Olympiad participants
frequently fail to solve three challenge problems in the
four-and-a-half hours allotted for that purpose, all four of the
collaborative MiniPolymath efforts generated a solution.
However, it should be noted that some of these solutions
took more than 24 hours to develop.
%%
IATC can help us understand how the proof efforts progressed, and can
potentially help us understand why they were (mathematically)
successful.
\end{itemize}

The plan of the work is as follows.
%%
\textsection\ref{background} reviews previous research on mathematical
argument, presents a brief introduction to Inference Anchoring Theory, and
describes Lamport's structured
proofs as an example of the state of the art for
modelling informal mathematical knowledge.
%%
\textsection\ref{iatc} introduces IATC, describes the grammar of IATC markup,
and describes the differences between this language and IAT.
%%
\textsection\ref{iatc-examples} presents our analysis of the examples
outlined above, which have been marked-up with IATC in order to
illustrate the relevant modelling concerns.
%%
\textsection\ref{discussion} summarises and reviews the contribution,
situates our work in relationship to the broader literature, and
outlines potential directions for further work.

@@

Our approach to argument builds on Buzynska and Reed's
Inference Anchoring Theory (IAT), which we describe in
Section \ref{iat}.  The specific conception of \emph{argument} that
underlies IAT is as follows:
\begin{quote}
[A]rguing can be interpreted as an illocutionary act that comes about as
the result of a relation between uttering a premise and uttering a
conclusion, thus mirroring the logical structure of inference[.]  \cite[p.~146]{Reed2017}
\end{quote}
\citet{reed2011dialogues} note that in everyday language the term
``argument'' is used to describe a particular kind of interaction
as well as the shared understanding extracted from these
interactions, as ``evidence'' or ``proof.''
%%   It is argument in the
%% latter sense which is \emph{reason-giving}: it ``expresses a point of
%% view on a question, and offers as support for this position one or
%% more reasons'' \cite[p.~448]{hitchcock2017informal}.
The purpose of IAT is to make the links between discourse and reasoning explicit.
%%

Concerning argumentation in a mathematical context, \citet[p.~39]{pedemonte2007} argues that
``analysis of the `content' is not sufficient to analyse all the cognitive aspects in
the relationship between argumentation and proof.''
%%
A large part of mathematical discussion is in essence meta-discussion
about meta-level objects, such as proof strategies that are
suggested on the fly and debates about whether these strategies are
likely to work as intended.
%%

\citet{mercier2011humans} distinguish
%% Reasoning ``enables communicators to produce arguments
%% to convince addressees who would not accept what they say on trust''
%% and, on the other hand, ``enables addressees to evaluate the soundness
%% of these arguments and to accept valuable information that they would
%% be suspicious of otherwise'' (p.~72).
arguments from inferences: 
only in the case of arguments ``the reasons for drawing this
conclusion on the basis of the premises are (at least partially)
spelled out'' (p.~58).   
%%
By contrast, formal mathematics is typically based on the reductive assumption that
``mathematical reasoning may be identified with classical, deductive
inference'' \cite[p.~25]{aliseda2003mathematical}.
However, everyday mathematical reasoning plainly involves more than just proof steps.
Here are two examples of familiar patterns of reasoning that appear in MiniPolymath 3:
\begin{description}
\item[\emph{argument from authority}]
``My bachelor thesis supervisor said that one can't use the word cardinal if we talk about finite sets. One has to use the words `number of elements'{''}
\cite[\href{https://wp.me/pAG2F-41\#comment-3418}{19 July, 9:46 pm}]{tao2011imo}.
\item[\emph{argument from analogy}]
``Let me check that I got the example correctly: is this `a point inside a regular polygon'? Isn't it established in an early comment that the example of a point inside an equilateral triangle indeed visits \emph{all} the points? Can you clarify the difference here?''
\cite[\href{https://wp.me/pAG2F-41\#comment-3378}{19 July, 9:19 pm}]{tao2011imo}.
\end{description}

%% Proofs are concerned with what necessarily follows from accepted premises
%% according to accepted inference rules.  There are, to be sure, other
%% relevant perspectives on mathematics, for example, the question of how we come to
%% accept or select certain premises and inference rules rather than
%% others.  Here, our concern will be communicated reasoning in
%% mathematical settings, looking at an evidence base that includes but
%% is not limited to deductive proofs.

@@

Some theorists have expressly contended that proofs are not arguments:
this is because proofs offer certainty, while arguments cannot
\cite{Dufour2013}.
Nevertheless, communication of reasons and reasoning
can be found throughout mathematical practice.
%% Aspects of
%% mathematical communication that go beyond
%% inference, and that support it, are particularly prevalent
%%
\citet{pedemonte2007} highlights the use of inductive and abductive logic
as well as deduction in mathematical processes that move from
``from conjecturing to the construction of proof [to] the proof as
product,'' and in which ``content rather than formal criteria''
can guide the proving process.  \citet{Dufour2013} gives examples
of argumentation ``not only before and during the proof but also after,
at least as long as it can be criticized'' (p.~74).
%% For example, is not appropriate to conclude
%% from $f(n)=f(-n)$ that $n=-n$ \cite[p.~141]{dove2009towards},
%% because functions and numbers are different kinds of mathematical content
%% with different associated rules.
%%
%% %\citet{pedemonte2007} disagrees and sees them as something more general.
%% %%
%% ``Argument'' may in some cases stand for the concept of an
%% , one hallmark being that
%% such presentations admit multiple semantically distinct formalisations
%% .  On the other hand, ``argument'' may stand
%% for a still more abstract 
%% which would not admit formalisation without the addition of
%% significant further details.  Yet again, in mathematics as in other
%% disciplines, the concept of ``argument'' is sometimes associated with
%% social phenomena of 
%%
Other scholars have observed features such as these:
%%
\begin{itemize}
%% The relationship between mathematical reasoning and everyday reasoning
%% has been debated \cite{kisielewicz2015new,aberdein2015commentary}.
\item Published mathematical writing tends to be particularly explicit about
reasons and conclusions \cite[p.~149]{dove2009towards}.
\item 
Not only the \emph{Prover} but also the \emph{Skeptic}
``has an important role to play, namely to ensure that the proof is persuasive, perspicuous, and valid'' \cite[p.~2618]{novaes2016reductio}.
%% \citet{aberdein2010virtue} 
%% has suggested that for argumentation to be successful,
%%  communication must sufficiently manifest several ``virtues of
%% argument'', broadly typologised as \emph{willingness to engage in
%%   argumentation}, \emph{willingness to listen to others},
%% \emph{willingness to modify one's own position}, and \emph{willingness
%%   to question the obvious}.
%%
\item On the way to a proof, degrees of
  confidence about the conclusions to be drawn may be discussed \cite[p.~17]{inglis2007modelling}.
%%
%%
\item Mathematical meanings need to be interpreted, and this tends to be a struggle  \cite[p.~360]{vanOers}.
\end{itemize}

\paragraph{\citet{Carrascal2015} provides an excellent survey of recent thinking about argument in mathematics, highlighting its connections with mathematical practice.}
Carrascal advises: ``in
order to learn more about the nature of mathematical practice and how
its products are evaluated, we should be looking at real examples of
this practice.''  She points to \citet{pease-and-martin} as a notable
example in this genre.  Once we have developed a suitable apparatus,
Section \ref{iatc-examples} will tackle several
real-world examples, including a detailed reexamination of the
dataset studied by Pease and Martin.
%% , which is sourced
%% from a public discussion of a problem that appeared in an advanced
%% high school level mathematical competition.

%% \paragraph{Carrascal emphasises that mathematical communication has a rhetorical aspect which depends on context.}
%% %%
%% %%

``Blog maths'' \cite{barany2010b} and other online discussions,
for example, on the
question-and-answer site MathOverflow, can ``tell us
about mathematicians' attitudes to working together in public'' as
well as the ``kinds of activities that go on in developing a proof''
\cite{Martin2015}.
%%
In the process of creating a proof or mathematical theory, divergent
understandings are negotiated using shared concepts, definitions, and
standards for proof, even as the concepts evolve.  Along these lines,
\citet{lak} used the methods of structured and abstract argumentation
to formalise the theory of informal mathematics developed in
Lakatos's \emph{Proofs and Refutations}
\citeyearpar{lakatos2015proofs} as a set of rules for turn-taking in a
dialogue game.  This work shows that formally specified and fully
implemented argumentation tools can be brought together and applied to
a specific, demanding, domain of human reasoning.\footnote{\label{fn:technical-details}The
  dialogue game defines ordered operations on a shared information state
  represented in the Argument Interchange Format (AIF)
  \cite{lawrence2012aifdb}, which is then interpreted by The Online
  Argument Structures Tool (TOAST) \cite{Snaith2012} and passed on to
  DungOMatic \cite{snaith2010pipelining} to calculate the grounded
  extension, which in this case represents the currently accepted,
  collaboratively constructed, proof or theory under discussion.}
%%
\citet{dauphin2018aspicend} produced a similar model of
natural-deduction style arguments, explanations,
and the ``\emph{prima facie} laws of logic'' such as may be debated in
work on mathematical foundations.
%%
These prior efforts focus on developing rules that give a plausible codification
of mathematical process.  Our concern is different,
but complementary.  We are interested in a better
understanding of what is actually said in mathematical arguments, and on
the reasoning that is conveyed.
%%
Accordingly, we will adapt a general-purpose argument modelling
approach, \emph{Inference Anchoring Theory}, which is described in the
following section.

@@

\paragraph{IAT is grounded in a notion of \emph{dialogical relations} that formalise the informal ``conventions and norms that dictate the flow of dialogue'' \cite{snaith2016dialogue}.}
Per \citet{budzynska2011speech}, these dialogical relations are also
referred to as ``\emph{transitions},'' a term that is meant to recall
the notion of transitions between operating states in a finite state
machine.  Indeed, when the norms have been fully codified in a
\emph{dialogue protocol}, the transitions are exactly described by a
finite state machine.\footnote{In such a setting the formal
  argumentation-theoretic techniques and tools mentioned in Footnote
  \ref {fn:technical-details} can be applied, though IAT models are
  not required to be fully formal in this regard.}  Content
relationships are typically identified by matching locutions against
known argument schemes, e.g., an `Argument from Positive Consequences'
is associated with two transitions, `challenging' and `substantiating'
\cite{walton08}.
%%
\citet{budzynska2014towards} describe Inference Anchoring Theory in terms of three components:
\begin{quote}
\begin{enumerate}[align=Center,label=(\roman*),labelsep=4pt,leftmargin=*,labelwidth=!]
\item\label{transitions-item} relations between locutions in a dialogue, called \emph{transitions};
\item\label{contents-item} relations between sentences (propositional \emph{contents} of
  locutions); and
\item\label{connections-item} \emph{illocutionary connections} that link locutions with their
  contents.
\end{enumerate}
\hfill \cite{budzynska2014towards}, emphasis added
\end{quote}

@@

Figure \ref{fig:TypicalIAT} is a typical example of an IAT analysis.
Figure \ref{fig:A-not-A} illustrates a feature
that was not directly mentioned in the list \ref{transitions-item}--\ref{connections-item}, above;
specifically,
this figure uses an \emph{{`}implicit{'} speech-act}
to anchor propositional content on a transition rather than a locution.
%%
Here, when a speaker asserts `$A$' and their interlocutor says
`\emph{No}', the logical content `$\neg A$' is
attached to the transition, rather than to the negating word.
The basic rationale is that the locution `\emph{No}' cannot be made sense
of without the preceding context.  There has been some debate about what to do about this.
%%
\citet{botting2015inferences} says that the choice to
anchor arguments on transitions is a conceptual mistake.  
However, for the creators of IAT, the reason illocutionary acts can
be rooted on dialogical relations follows
\begin{quote}
\ldots directly from pragma-dialectical analysis which views
the speech act of assertion [\lips] as occurring at the `sentence'
level, and the speech act of argumentation as occurring at a `higher
textual level.' \cite{budzynska2011speech}
\end{quote}

@@

\citet{visser2011correspondence} describe the theoretical
considerations in more detail.
The pattern common to both Figure \ref{fig:TypicalIAT} and Figure
\ref{fig:A-not-A} is that allowable inferences are governed by
dialogue norms.  In Figure \ref{fig:TypicalIAT}, for instance, we
would not immediately know that `$A^\prime$' is intended to \emph{support} `$A$'
without Wilma's intermediate question which explicitly requested such
support.  Given the context, the intended
inference is is clear.
%%
Thus, both examples serve to illustrate that
\begin{quote}
the connection between locutions in a dialogue has an inferential
component beyond any that may hold between the contents of those
locutions \cite{reed2011dialogues}.
\end{quote}
%%
In short, IAT studies ``the way in which the rules of dialogue
influence the construction of argument''
\cite{budzynska2016theoretical}.

@@

For now, we comment that while the use of strict hierarchies
is not representative of the way proofs are usually
constructed in day-to-day practice, Lamport has proposed that
structured proofs can assist in proof development, e.g., by helping to
bring errors to the surface.  However, they do not necessarily make the job of
the reader easier: Lamport \citeyearpar[p.~20]{lamport2012write21st} quotes a
referee who had read one of his structured proofs:
\begin{quote}
The proofs [\lips] are lengthy, and are presented in a style which I
find very tedious.  [\lips] My feeling is that informal proof sketches
[\lips] to explain the crucial ideas in each result would be more
appropriate.
\end{quote}
%%
Unlike structured proofs, IATC is intended to express the typical processes
by which proofs are generated in standard practice, rather than
make the process of proving and reading proofs easier.
It would nevertheless be compatible with our
aims to include formal statements in {TLA}$^{+}$ (or some other
language) in IATC's content layer.

@@

For example, the long-range \texttt{reform} connection from $A$ to $A^\prime$ in
our content analysis would suggest a corresponding long-range transition from
Bob's first to his last statement in the dialogue.
However, that would still neglect Bob's so-far implicit reasoning
to the effect that $A^\prime$ is (potentially) not vulnerable to objection $B$.
If the dialogue continued from this point, detailed relationships between the constituent contents of `$A^\prime$' and `$B$' may need to be discussed, and an IATC analysis would be able to unpack these and account for the details.
%%   For example, $A^\prime$ relates explicitly to the content of
%% Bob's original statement ``$A$'', and is modelled as a reformation of that earlier
%% statement.  This suggests a long-range transition between Bob's first and second locutions.  Our current considerations

>>>>

@@

\begin{abstract}
To adequately model mathematical arguments the analyst must be able to
represent the mathematical objects under discussion and the
relationships between them, as well as inferences drawn about these
objects and relationships as the discourse unfolds.
We introduce a framework with these properties, which has been 
used to analyse mathematical dialogues and expository texts.
The framework can recover salient elements of discourse at, and within, the sentence
level, as well as the way mathematical content connects to form larger
argumentative structures.  We show how the framework might be used to
support computational reasoning, and argue that it provides a more
natural way to examine the process of proving theorems than do
Lamport's structured proofs.
\keywords{inference anchoring theory \and mathematical practice \and mathematical argument \and structured proof}
\end{abstract}

@@

IATC could be overlaid upon formally specified contents, where these are available.  Lamport's ``Temporal Logic of Actions+'' (TLA+) \cite{lamport1999specifying,lamport2014tla2} is one such formalism that could be used to model content-level expressions.  Higher-level discourse structure would then be exhibited somewhat along the lines of Lamport's own semi-formal ``structured proofs'' 
\cite{lamport1995write,lamport2012write21st}.  However, unlike structured proof, IATC does not aim to reshape the way people do mathematics, but to model it more exactly.  As such, it constitutes groundwork for a future generation of computer systems that can collaborate with mathematicians and students in a way these potential users already understand. \citet{epstein2015wanted} highlights the ``extent to which a person believes that her work experience or product has been facilitated or improved by the collaboration'' as a key evaluation metric for assessing collaborative intelligent computer systems. The key metric at this stage is more basic, namely, we are interested in the degree to which IATC can represent real-world examples of mathematical practice in a way that can make them accessible to computational reasoning. After introducing the modelling approach, we use several examples to show that IATC is indeed satisfactory in this regard.

\begin{itemize}
\item Our first example is a school-level challenge problem that was
  presented in a public lecture by the mathematician Timothy Gowers
  \cite{gowers-talk}.  The lecture aimed to motivate and contextualise
  a project, then beginning, to develop mathematical software that
  ``operate[s] in a way that closely mirrors the way human
  mathematicians operate'' \cite[p.~255]{ganesalingam2016fully}.  The
  reasoning needed to solve the challenge problem remains beyond the
  scope of the computational method that Ganesalingam and Gowers
  ultimately published, but it is both sufficiently simple and
  sufficiently realistic to introduce the practical aspects of
  working with IATC.
\item Our second example is a question posed on the online Q\&A forum
  MathOverflow, together with the ensuing dialogue.  MathOverflow is
  part of the Stack Exchange network of community question-and-answer
  websites, which is particularly popular with software developers.
  The MathOverflow sub-site is devoted to discussions about
  research-level questions in mathematics.  Such discussions are very
  different from the textbook-style proofs treated by
  \cite{ganesalingam2016fully}, and we discuss the considerations that
  such discussions would impose on computational modelling efforts.
\item \paragraph{MiniPolymath 1 through 4 were part of a series of
  experiments in collaborative online mathematics known as ``Polymath
  projects'' \cite{polymath-wiki}.}  While other projects in the series
  tackled novel research, the problems in the MiniPolymath subseries
  were drawn from the Mathematical Olympiad, a premier competition for
  pre-college students.
Six problems are given, and the examination takes place over
two days with three problems to be solved each day.
Whereas individual Olympiad participants
frequently fail to solve three challenge problems in the
four-and-a-half hours allotted for that purpose, all four of the
collaborative MiniPolymath efforts generated a solution.
However, it should be noted that some of these solutions
took more than 24 hours to develop.
IATC can help us understand how the proof efforts progressed, and can
potentially help us understand why they were (mathematically)
successful.
\end{itemize}

The plan of the work is as follows.
\textsection\ref{background} reviews previous research on mathematical
argument, presents a brief introduction to Inference Anchoring Theory, and
describes Lamport's structured
proofs as an example of the state of the art for
modelling informal mathematical knowledge.
\textsection\ref{iatc} introduces IATC, describes the grammar of IATC markup,
and describes the differences between this language and IAT.
\textsection\ref{iatc-examples} presents our analysis of the examples
outlined above, which have been marked-up with IATC in order to
illustrate the relevant modelling concerns.
\textsection\ref{discussion} summarises and reviews the contribution,
situates our work in relationship to the broader literature, and
outlines potential directions for further work.

@@

Our approach to argument builds on Buzynska and Reed's
Inference Anchoring Theory (IAT), which we describe in
Section \ref{iat}.  The specific conception of \emph{argument} that
underlies IAT is as follows:
\begin{quote}
[A]rguing can be interpreted as an illocutionary act that comes about as
the result of a relation between uttering a premise and uttering a
conclusion, thus mirroring the logical structure of inference[.]  \cite[p.~146]{Reed2017}
\end{quote}
\citet{reed2011dialogues} note that in everyday language the term
``argument'' is used to describe a particular kind of interaction
as well as the shared understanding extracted from these
interactions, as ``evidence'' or ``proof.''
The purpose of IAT is to make the links between discourse and reasoning explicit.

Concerning argumentation in a mathematical context, \citet[p.~39]{pedemonte2007} argues that
``analysis of the `content' is not sufficient to analyse all the cognitive aspects in
the relationship between argumentation and proof.''
A large part of mathematical discussion is in essence meta-discussion
about meta-level objects, such as proof strategies that are
suggested on the fly and debates about whether these strategies are
likely to work as intended.

\citet{mercier2011humans} distinguish
arguments from inferences: 
only in the case of arguments ``the reasons for drawing this
conclusion on the basis of the premises are (at least partially)
spelled out'' (p.~58).   
By contrast, formal mathematics is typically based on the reductive assumption that
``mathematical reasoning may be identified with classical, deductive
inference'' \cite[p.~25]{aliseda2003mathematical}.
However, everyday mathematical reasoning plainly involves more than just proof steps.
Here are two examples of familiar patterns of reasoning that appear in MiniPolymath 3:
\begin{description}
\item[\emph{argument from authority}]
``My bachelor thesis supervisor said that one can't use the word cardinal if we talk about finite sets. One has to use the words `number of elements'{''}
\cite[\href{https://wp.me/pAG2F-41\#comment-3418}{19 July, 9:46 pm}]{tao2011imo}.
\item[\emph{argument from analogy}]
``Let me check that I got the example correctly: is this `a point inside a regular polygon'? Isn't it established in an early comment that the example of a point inside an equilateral triangle indeed visits \emph{all} the points? Can you clarify the difference here?''
\cite[\href{https://wp.me/pAG2F-41\#comment-3378}{19 July, 9:19 pm}]{tao2011imo}.
\end{description}

@@

Some theorists have expressly contended that proofs are not arguments:
this is because proofs offer certainty, while arguments cannot
\cite{Dufour2013}.
Nevertheless, communication of reasons and reasoning
can be found throughout mathematical practice.
\citet{pedemonte2007} highlights the use of inductive and abductive logic
as well as deduction in mathematical processes that move from
``from conjecturing to the construction of proof [to] the proof as
product,'' and in which ``content rather than formal criteria''
can guide the proving process.  \citet{Dufour2013} gives examples
of argumentation ``not only before and during the proof but also after,
at least as long as it can be criticized'' (p.~74).
Other scholars have observed features such as these:
\begin{itemize}
\item Published mathematical writing tends to be particularly explicit about
reasons and conclusions \cite[p.~149]{dove2009towards}.
\item 
Not only the \emph{Prover} but also the \emph{Skeptic}
``has an important role to play, namely to ensure that the proof is persuasive, perspicuous, and valid'' \cite[p.~2618]{novaes2016reductio}.
\item On the way to a proof, degrees of
  confidence about the conclusions to be drawn may be discussed \cite[p.~17]{inglis2007modelling}.
\item Mathematical meanings need to be interpreted, and this tends to be a struggle  \cite[p.~360]{vanOers}.
\end{itemize}

\paragraph{\citet{Carrascal2015} provides an excellent survey of recent thinking about argument in mathematics, highlighting its connections with mathematical practice.}
Carrascal advises: ``in
order to learn more about the nature of mathematical practice and how
its products are evaluated, we should be looking at real examples of
this practice.''  She points to \citet{pease-and-martin} as a notable
example in this genre.  Once we have developed a suitable apparatus,
Section \ref{iatc-examples} will tackle several
real-world examples, including a detailed reexamination of the
dataset studied by Pease and Martin.

``Blog maths'' \cite{barany2010b} and other online discussions,
for example, on the
question-and-answer site MathOverflow, can ``tell us
about mathematicians' attitudes to working together in public'' as
well as the ``kinds of activities that go on in developing a proof''
\cite{Martin2015}.
In the process of creating a proof or mathematical theory, divergent
understandings are negotiated using shared concepts, definitions, and
standards for proof, even as the concepts evolve.  Along these lines,
\citet{lak} used the methods of structured and abstract argumentation
to formalise the theory of informal mathematics developed in
Lakatos's \emph{Proofs and Refutations}
\citeyearpar{lakatos2015proofs} as a set of rules for turn-taking in a
dialogue game.  This work shows that formally specified and fully
implemented argumentation tools can be brought together and applied to
a specific, demanding, domain of human reasoning.\footnote{\label{fn:technical-details}The
  dialogue game defines ordered operations on a shared information state
  represented in the Argument Interchange Format (AIF)
  \cite{lawrence2012aifdb}, which is then interpreted by The Online
  Argument Structures Tool (TOAST) \cite{Snaith2012} and passed on to
  DungOMatic \cite{snaith2010pipelining} to calculate the grounded
  extension, which in this case represents the currently accepted,
  collaboratively constructed, proof or theory under discussion.}
\citet{dauphin2018aspicend} produced a similar model of
natural-deduction style arguments, explanations,
and the ``\emph{prima facie} laws of logic'' such as may be debated in
work on mathematical foundations.
These prior efforts focus on developing rules that give a plausible codification
of mathematical process.  Our concern is different,
but complementary.  We are interested in a better
understanding of what is actually said in mathematical arguments, and on
the reasoning that is conveyed.
Accordingly, we will adapt a general-purpose argument modelling
approach, \emph{Inference Anchoring Theory}, which is described in the
following section.

@@

\paragraph{IAT is grounded in a notion of \emph{dialogical relations} that formalise the informal ``conventions and norms that dictate the flow of dialogue'' \cite{snaith2016dialogue}.}
Per \citet{budzynska2011speech}, these dialogical relations are also
referred to as ``\emph{transitions},'' a term that is meant to recall
the notion of transitions between operating states in a finite state
machine.  Indeed, when the norms have been fully codified in a
\emph{dialogue protocol}, the transitions are exactly described by a
finite state machine.\footnote{In such a setting the formal
  argumentation-theoretic techniques and tools mentioned in Footnote
  \ref {fn:technical-details} can be applied, though IAT models are
  not required to be fully formal in this regard.}  Content
relationships are typically identified by matching locutions against
known argument schemes, e.g., an `Argument from Positive Consequences'
is associated with two transitions, `challenging' and `substantiating'
\cite{walton08}.
\citet{budzynska2014towards} describe Inference Anchoring Theory in terms of three components:
\begin{quote}
\begin{enumerate}[align=Center,label=(\roman*),labelsep=4pt,leftmargin=*,labelwidth=!]
\item\label{transitions-item} relations between locutions in a dialogue, called \emph{transitions};
\item\label{contents-item} relations between sentences (propositional \emph{contents} of
  locutions); and
\item\label{connections-item} \emph{illocutionary connections} that link locutions with their
  contents.
\end{enumerate}
\hfill \cite{budzynska2014towards}, emphasis added
\end{quote}

@@

Figure \ref{fig:TypicalIAT} is a typical example of an IAT analysis.
Figure \ref{fig:A-not-A} illustrates a feature
that was not directly mentioned in the list \ref{transitions-item}--\ref{connections-item}, above;
specifically,
this figure uses an \emph{{`}implicit{'} speech-act}
to anchor propositional content on a transition rather than a locution.
Here, when a speaker asserts `$A$' and their interlocutor says
`\emph{No}', the logical content `$\neg A$' is
attached to the transition, rather than to the negating word.
The basic rationale is that the locution `\emph{No}' cannot be made sense
of without the preceding context.  There has been some debate about what to do about this.
\citet{botting2015inferences} says that the choice to
anchor arguments on transitions is a conceptual mistake.  
However, for the creators of IAT, the reason illocutionary acts can
be rooted on dialogical relations follows
\begin{quote}
\ldots directly from pragma-dialectical analysis which views
the speech act of assertion [\lips] as occurring at the `sentence'
level, and the speech act of argumentation as occurring at a `higher
textual level.' \cite{budzynska2011speech}
\end{quote}

@@

\citet{visser2011correspondence} describe the theoretical
considerations in more detail.
The pattern common to both Figure \ref{fig:TypicalIAT} and Figure
\ref{fig:A-not-A} is that allowable inferences are governed by
dialogue norms.  In Figure \ref{fig:TypicalIAT}, for instance, we
would not immediately know that `$A^\prime$' is intended to \emph{support} `$A$'
without Wilma's intermediate question which explicitly requested such
support.  Given the context, the intended
inference is is clear.
Thus, both examples serve to illustrate that
\begin{quote}
the connection between locutions in a dialogue has an inferential
component beyond any that may hold between the contents of those
locutions \cite{reed2011dialogues}.
\end{quote}
In short, IAT studies ``the way in which the rules of dialogue
influence the construction of argument''
\cite{budzynska2016theoretical}.

@@

For now, we comment that while the use of strict hierarchies
is not representative of the way proofs are usually
constructed in day-to-day practice, Lamport has proposed that
structured proofs can assist in proof development, e.g., by helping to
bring errors to the surface.  However, they do not necessarily make the job of
the reader easier: Lamport \citeyearpar[p.~20]{lamport2012write21st} quotes a
referee who had read one of his structured proofs:
\begin{quote}
The proofs [\lips] are lengthy, and are presented in a style which I
find very tedious.  [\lips] My feeling is that informal proof sketches
[\lips] to explain the crucial ideas in each result would be more
appropriate.
\end{quote}
Unlike structured proofs, IATC is intended to express the typical processes
by which proofs are generated in standard practice, rather than
make the process of proving and reading proofs easier.
It would nevertheless be compatible with our
aims to include formal statements in {TLA}$^{+}$ (or some other
language) in IATC's content layer.

@@

For example, the long-range \texttt{reform} connection from $A$ to $A^\prime$ in
our content analysis would suggest a corresponding long-range transition from
Bob's first to his last statement in the dialogue.
However, that would still neglect Bob's so-far implicit reasoning
to the effect that $A^\prime$ is (potentially) not vulnerable to objection $B$.
If the dialogue continued from this point, detailed relationships between the constituent contents of `$A^\prime$' and `$B$' may need to be discussed, and an IATC analysis would be able to unpack these and account for the details.
